---
title: "linear model from Web"
author: "Destiny Calvillo"
date: "Summer 2021"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

 


```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

## Model 1 SIM~ARMS

### scatterplot

```{r}
basicNN <- ggplot(data,aes(y=SIMS,x=ARM))
```
```{r}
basicNN + geom_point()
```

The Y-axis is SIMS and the X-axis is ARM. This plot shows that there is a positive linear correlation. The plot above displays a positive correlation between ARM and SIMS.

### Numerical results

```{r}
cor(SIMS~ARM,data=data)
```

Above is the value of correlation, the value is 0.68. Since the value is a positive number, we should expect the sea of points to go uphill towards the right. We can also verify this in the scatter-plots from earlier.

### Inferential  (Build model.1)

```{r}
model.1 <- lm(SIMS~ARM,data=data)
summary.lm(model.1)
```

For the inferential results, we can see that the Adjusted R-Square is bigger, which means that more errors were reduced. To calculate SIMS based on ARM strength, we would use the equation of a line. The equation that best fits this is SIMS= -4.1 + 0.05 * R. The formula needed is y= mx + b to create the corresponding line. The Adjusted R-squared is the number to look at, which shows that the error was reduced by an estimated 47% or 0.467. Also, the prediction interval is between 1.226. There is a 145 degrees of freedom, which indicates the amount of data points or dots that there are in the plot.

#### Predict at target point

```{r}
newdata = data.frame(GRIP = 94, ARM = 88)
predict(model.1, newdata, interval = "prediction")
```

Above is the prediction interval given the targeted points. Meaning that it shows how far one point varies from the other. Furthermore, we'd end up with an approx amount of 2.3 if we were to subtract the upper and fit. 

#### scatterplot with model fit
 
```{r}
basicNN + geom_point( ) + geom_smooth(method=lm)
```   

Just by simply looking at the plot, it appears that the value of the Y-intercept would be around 1 or 1.25. The dots presented are apart of the regression line which was formed using the formula of a line. To calculate SIMS based on ARM strength, we would use the equation of a line. The equation that best fits this is SIMS= -4.1 + 0.05 * R. The formula needed is y= mx + b to create the corresponding line. Each point represents an output of the standard deviation within the normal distribution.

## Model 2 SIM~GRIP

### Now add in scatterplot


### Numerical results 


### Inferential  (Build model.2)
  

#### predict model.2 at target point

#### now add the model fit to our plot for model.2
  


## Model 3 SIM~ARM+GRIP

### Numerical results (cor)

`
  
### Inferential  (Build 2-dimentional model.3)

#### predict model.3 at target point
  

## Comparing nested models ANOVA Test

### Model.1 vs Model.3
### Model.2 vs Model.3

## Informally compare Model.1 with model.2

